---
title: "spike-regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spike-regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Rexpneuro)
suppressPackageStartupMessages(library(tidyverse))
library(broom)
library(future)
library(tictoc)

future::plan("multisession")
```

Read data into memory,
```{r echo = TRUE, message = FALSE, warning = FALSE}
basedir <- "~/ownCloud/behaviordata/data_analyses_Brian"

# Read in eventide and corresponding spike data, should take ~30 seconds
tic()
obj <- read_eventide(name = c("flocky", "tess"), include_tracker = F, basedir = basedir) %>%
  read_matched_spike_data(basedir = basedir)
toc()

# Brief summary
summary(obj)

# total number of neurons
sum(obj$info %>% mutate(n = map_dbl(neuron_info, ~nrow(.x))) %>% pull(n))

# Glance at time differences between plexon and EventIDE
skim(obj$trial_duration %>% select(session, dt))
```

The data returned is not yet useful for running regressions. Let's create a more
useful data frame,
```{r echo = TRUE, message = FALSE, warning = FALSE}
tic()
# Takes about 8 seconds, the align parameter must exist as a column in obj$trial_data
df <- prep_for_model(obj, align = "target_onset_time", t_start = 0, t_end = 0.5, min_trial = 50)
toc()
```

The call above defaults to binning spikes into a single window, and we can now run
it through a linear regression (per neuron),
```{r}
t1 <- Sys.time()
res <- df %>% group_by(uname, t) %>%
  group_split() %>%
  map_dfr(.f = function(df) {
    lm(fr ~ block*direction + gng*direction, data = df) %>%
      tidy() %>%
      add_column(t = unique(df$t), .before = 1) %>%
      add_column(uname = unique(df$uname), .before = 1) %>%
      add_column(session = unique(df$session), .before = 1)
  }) %>%
  arrange(session)
Sys.time() - t1
```
The data frame returned contains a separate row per coefficient, per neuron
```{r fig.width = 6, fig.height = 4, fig.align = "center"}
res

# Filter by coefficient and pipe to figure
res %>% 
  filter(term == "gnggo") %>% 
  mutate(thresh = p.value < 0.05) %>% 
  ggplot(aes(estimate, fill = thresh)) + geom_histogram()
```

If we wanted to separately plot histograms of coefficients for neurons that 
increased or decreased their response (average over trials), we need to go back
to the prepped data frame,

```{r fig.width = 6, fig.height = 4, fig.align = "center"}
# Determine whether neuron is positive-response relative to baseline
temp = df %>% 
  group_by(uname) %>%                       # for each unique neuron
  mutate(fr_mean = mean(fr)) %>%            # calculate mean firing rate in window
  mutate(pos = fr_mean > fr_bl_mean) %>%    # compare to baseline
  slice(1) %>%                              # take first row
  arrange(session)

res %>% left_join(temp %>% select(id, session, uname, pos)) %>%
  filter(term == "gnggo") %>% 
  mutate(thresh = p.value < 0.05) %>% 
  ggplot(aes(estimate, fill = thresh)) + geom_histogram() + facet_wrap(~pos)
```

We can also run multiple timebins using the same functions,
```{r fig.width = 6, fig.height = 4, fig.align = "center"}
# Note the binwidth parameter
df <- prep_for_model(obj, align = "target_onset_time", t_start = -0.5, t_end = 0.5, 
                     binwidth = 0.1, min_trial = 50)

t1 <- Sys.time()
res <- df %>% group_by(uname, t) %>%
  group_split() %>%
  map_dfr(.f = function(df) {
    lm(fr_norm3 ~ block*direction + gng*direction, data = df) %>%
      tidy() %>%
      add_column(t = unique(df$t), .before = 1) %>%
      add_column(uname = unique(df$uname), .before = 1) %>%
      add_column(session = unique(df$session), .before = 1)
  }) %>%
  arrange(session)
Sys.time() - t1
```

We could use a separate plot for each time bin of a specific term, although this 
is a little busy,
```{r fig.width = 7, fig.height = 7, fig.align = "center"}
res %>% 
  filter(term == "blockmix") %>% 
  mutate(thresh = p.value<0.05) %>% 
  ggplot(aes(estimate, fill=thresh)) + geom_histogram() + facet_wrap(~as.factor(t))
```

Or more compactly represent all the terms,
```{r fig.width = 8, fig.height = 8, fig.align = "center"}
library(ggridges)
res %>% 
  mutate(thresh = p.value<0.05) %>% 
  ggplot(aes(x = estimate, y = as.factor(t), fill=thresh)) + 
  geom_density_ridges2(stat = "binline", alpha = 0.5, scale = 0.95, bins = 60) + 
  facet_wrap(~term)

# If we want to restrict to a particular brain area
res %>% left_join(obj$info %>% filter(target=="gpi")) %>% filter(target == "gpi") %>%
  mutate(thresh = p.value<0.05) %>% 
  ggplot(aes(x = estimate, y = as.factor(t), fill=thresh)) + 
  geom_density_ridges2(stat = "binline", alpha = 0.5, scale = 0.95, bins = 60) + 
  facet_wrap(~term)
```
